\propertitle

### 2019-06-25 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-25T17:30:45+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The method introduced in the present work reveals an issue in the recipe
above. The issue doesn't come from the method itself but from a simple,
general fact of probability theory: if a probability distribution has some
sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
then its marginals \emph{cannot} have the same sufficient statistics, and
vice versa (except for trivial cases, such as uniform distributions). This
impossibility is known in statistical mechanics: marginals of Gibbs states
aren't Gibbs states \citep{maesetal1999}. Mathematically it corresponds to
the general impossibility of solving a system of $n$ equations in $m$
unknowns when $n>m$
\citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.

In the neuroscientific application above, this means for example that if
pairwise correlations are sufficient statistics for a sample of a network,
then they \emph{cannot} be sufficient for the full network; and vice versa,
if we assume that they are sufficient statistics for the full network, then
we must expect them \emph{not} to be sufficient for any recorded sample
thereof. Thus, maximum-entropy methods applied at the sample level
paradoxically give an answer opposite to the one we might think they're
giving!

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-25T17:40:21+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The method introduced in the present work reveals an issue in the recipe
above. The issue doesn't come from the method itself but from a simple,
general fact of probability theory: if a probability distribution has some
sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
then its marginals \emph{cannot} have the same sufficient statistics, and
vice versa (except for trivial cases, such as uniform distributions). This
also holds among marginals of different order (to $n$ neurons or to $n+1$
neurons, for example). This impossibility is known in statistical
mechanics: marginals of Gibbs states aren't Gibbs states
\citep{maesetal1999}. Mathematically it corresponds to the general
impossibility of solving a system of $n$ equations in $m$ unknowns when
$n>m$ \citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.

In the neuroscientific application above, this means for example that if
pairwise correlations are sufficient statistics for a sample of a network,
then they \emph{cannot} be sufficient for the full network or for a sample
with more or fewer neurons; and vice versa, if we assume that they are
sufficient statistics for the full network, then we must expect them
\emph{not} to be sufficient for any recorded sample thereof.

It seems to us that such questions about sufficient statistics are
implicitly asked for specific brain areas, not for casually recorded
samples thereof. Thus, maximum-entropy methods applied at the sample level
paradoxically give an answer opposite to the one we might think they're
giving!
